A deeper look into bug fixes: Patterns, replacements, deletions, and additions

### 2016-02-19 MSRChallenge2016.tex
characteristics, related work

We begin by broadly characterizing the bug-fixing changes to Java code in our
dataset.  We perform our analysis on the \emph{2015 September/Github} dataset on
Boa, which includes $7,830,023$ Java projects with $23,229,406$ revisions. Among
all the revisions, $4,590,679$ revisions are identified by Boa as bug fixing
revisions.

\subsubsection*{How many files are changed to fix a bug?} \vspace{-0.3em}

A file was changed if (1) it was a new file; (2) or it was modified; (3) or it
was deleted. In total, $52,052,571$ files got changed. Thus, on average each bug
fixing revision changes $11.3$ files.  This number is surprisingly high because
most automatic program repair techniques assume that bugs are local in most
cases. There could be several reasons behind this number. It could be the case
that other software artifacts (e.g.  documentation) are updated after bug
fixing. It could also be the case that bug fixing commit also contains changes
that are not related to fixing bug (e.g. refactoring, adding new features by
creating new files). To better understand why so many files are changed, we
continue to ask:

\subsubsection*{What are the file types of those changed files?} \vspace{-0.3em}

To answer this question, we iterate all the changed files in each bug fixing
revision and count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

\begin{table}
\centering
  \begin{tabular}{| c | r | r |}
  \hline
  File Kind & Total Number & Average \\ \hline \hline
  BINARY & 752,945 & 0.16 \\ \hline
  SOURCE\_JAVA\_ERROR & 2,073,558 & 0.45 \\ \hline
  SOURCE\_JAVA\_JLS2 & 2,607,413 & 0.57 \\ \hline
  SOURCE\_JAVA\_JLS3 & 15,748,967 & 3.43 \\ \hline
  SOURCE\_JAVA\_JLS4 & 83,798 & 0.02 \\ \hline
  TEXT & 541,023 & 0.12 \\ \hline
  XML & 6,818,299 & 1.49 \\ \hline
  UNKNOWN & 23,426,568 & 5.10 \\ \hline
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}

SOURCE\_JAVA\_ERROR file kind represents files that Boa failed to parse. JLS2,
JLS3, and JLS4 represent Java files written in different J2SE version. Boa
documentation does not explicitly explain what is TEXT and what is UNKNOWN. From
Table~\ref{tbl:fileType}, we can see that TEXT files and BINARY files are
changed least frequently. This is not surprising because TEXT files are usually
documentation files, and BINARY files should only be changed in rare case (e.g.
milestone, new release, dependency update). If we consider all 4 kinds of Java
files as one kind, then on average each bug fixing revision changes $4.47$ Java
files. But still, it does not conform to our assumption of bug locality. XML
files in Java projects usually represents build files, thus changing one build
file after each bug fixing is not surprising.  What surprises us is that UNKNOWN
files are changed most frequently. Note that all the projects in our dataset are
Java projects, so it is reasonable to assume that these files are not source
files that written in other programming languages.  One possible explanation is
that they are test resources. However, due to the limitations of Boa, we can not
look into the content of these files.

It could be the case that among those $4.47$ changed Java files, some of them
are related to testing, such as modifying the test cases and creating new test
classes. To better understand what changes are usually made to Java source code
files, we continue to ask:

\subsubsection*{How often do developers introduce new classes, methods, fields,
	local variables in
bug fixing commits?} \vspace{-0.3em}

This question is interesting because most automatic program repair approaches
often do not create new classes/methods/fields/local variables to fix bugs.
Most approaches guess and synthesize the patches based on existing source code,
and direct translations of approaches from C to Java do not naturally consider
object-oriented concerns, like fields or classes.

Table~\ref{tbl:new} shows our analysis results. It shows that introducing new
classes and new variables are rare, which is good news for automatic program
repair approaches. On average, we still have $0.69$ new method per bug fix. If
we consider the possibility of creating new test method (i.e.  methods that
annotated by @Test if JUnit is used), this number is not that scary. However,
what surprises us is that $1.32$ fields are created per bug fix.  This indicates
that automatic program repair approaches should pay attention to the state of
the class when fixing Java programs.

\begin{table}
\centering
  \begin{tabular}{| c | r | r |}
  \hline
  Introducing & Total Number & Average \\ \hline \hline
  Class & 729,201 & 0.16 \\ \hline
  Methods & 3,186,867 & 0.69 \\ \hline
  Fields & 6,076,646 & 1.32 \\ \hline
  Variables & 924,259 & 0.2 \\ \hline
  \end{tabular}
  \caption{Introducing new classes, methods, fields, local variables to fix bugs}
  \label{tbl:new}
\end{table}

@@

This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,weimer2009,legoues2012} first
find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced. With this study, we are able to make a more accurate guess based
on the phenomenon we observed from Table~\ref{tab:likeliness} and
Figure~\ref{fig:StmtType}.

>>>>

We begin by broadly characterizing the bug-fixing changes to Java code in the
dataset.  We perform our analysis on the \emph{2015 September/Github} dataset on
Boa, including $7,830,023$ Java projects with $23,229,406$ revisions. Boa
identifies $4,590,679$ as bug fixing.

\vspace{1ex}
\noindent\textbf{How many files are changed to fix a bug?}
%
We consider a file changed if it is new, modified, or
deleted in a commit. In total, $52,052,571$ files were changed. Thus, on average each bug
fixing revision changes $11.3$ files.  This is surprisingly high:
most automatic program repair techniques assume that bugs are local.
We hypothesize several explanations. Other
software artifacts (e.g.  documentation) may be updated after bug  
fixing. It could also be the case that bug fixing commit also contains changes
that are not related to fixing bug (e.g. refactoring, adding new features by
creating new files). To better understand this
phenomenon, we ask:

\vspace{1ex}
\noindent\textbf{What are the file types of those changed files?}
%
To answer this question, we iterate over all changed files in each bug fixing
revision and count the total number for each file type. Results are shown in
Table~\ref{tbl:fileType}.

\begin{table}
\centering
  \begin{tabular}{ l | r  r }
  \toprule
  File Kind & Total & Average \\ 
  \midrule
  Binary & 752,945 & 0.16 \\ 
  Java (ERROR) & 2,073,558 & 0.45 \\ 
  Java (JLS2) & 2,607,413 & 0.57 \\ 
  Java (JLS3) & 15,748,967 & 3.43 \\
  Java (JLS4)  & 83,798 & 0.02 \\ 
  Text & 541,023 & 0.12 \\ 
  XML & 6,818,299 & 1.49 \\ 
  UNKNOWN & 23,426,568 & 5.10 \\ 
  \end{tabular}
  \caption{File Types of Changed Files}
  \label{tbl:fileType}
\end{table}

JAVA (ERROR) files are those that Boa failed to parse. JLS2,
JLS3, and JLS4 represent Java files written in different J2SE versions. From
Table~\ref{tbl:fileType}, we see that Text and Binary files are
changed least frequently. This is unsurprising, since text are often
documentation files, and binaries should only be changed rarely. If we consider all 4 kinds of Java
files as one kind, on average, each bug fixing revision changes $4.47$ Java
files. This still does not conform to our assumption of bug locality. XML
files in Java projects usually represents build files; changing a build
file is unsurprising.  Rather more surprising is how frequently UNKNOWN files
are changed. Given that all the projects are Java projects, 
it is unclear whether these are written in other proramming
languages; One possibility is
that they are test resources. However, due to the limitations of Boa, we can not
look into the content of these files. 
\todo{Possibly mention median.} 
To better understand what changes are usually made to Java source code
files, we next ask:

\vspace{1ex}
\noindent\textbf{How often do developers introduce new classes, methods, fields,
	local variables in
bug fixing commits?} 
%
This question is interesting because most automatic program repair approaches
often do not create new methods or variables to fix bugs in C.
Most approaches guess and synthesize the patches based on existing source code,
and direct translations of approaches from C to Java do not naturally consider
object-oriented concerns, like fields or classes.

Table~\ref{tbl:new} shows that introducing new
classes and new variables is rare. On average, we still have $0.69$ new methods
per bug fix. These methods naturally contain new test methods (annotated by
@Test, for example), which is not too alarming. However,
what surprises us is that $1.32$ fields are created per bug fix.  This indicates
that automatic program repair approaches should pay attention to the state of
the class when fixing Java programs.

\begin{table}
\centering
  \begin{tabular}{l | r  r }
  \toprule
  Introducing & Total Number & Average \\ \midrule
  Class & 729,201 & 0.16 \\ 
  Methods & 3,186,867 & 0.69 \\ 
  Fields & 6,076,646 & 1.32 \\ 
  Variables & 924,259 & 0.2 \\ 
  \end{tabular}
  \caption{Introducing new classes, methods, fields, local variables to fix bugs}
  \label{tbl:new}
\end{table}

@@

This study is directly applicable to automatic software repair. Most common
tools for automatic software repair~\cite{kim2013,legoues2012} first
find a faulty statement using several different fault localization
techniques~\cite{fry2010}, and once they have that faulty statement, they apply
different kinds of edits to it, usually at random. Which makes it equally likely
to replace an statement for another without looking at the kinds of statements
being replaced. With this study, we are able to make a more accurate guess based
on the phenomenon we observed from Table~\ref{tab:likeliness} and
Figure~\ref{fig:StmtType}.
